Grounding the Interaction: Anchoring Situated Discourse in Everyday Human-Robot Interaction

### 2011-10-21 paper_knowledge.tex
Raquel contribution to try to cut down the number of pages

In this work we propose an approach aiming at grounding the interaction with
users. Figure~\ref{fig|overall-process} illustrates the general contex. We have
developed three distinct, inter-related cognitive components, as illustrated in
Figure~\ref{fig|main-components}.

@@

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.8\linewidth]{images/grounding_robot.png}
  \caption{Generic model of cognitive abilities interactions for grounding}
  \label{fig|overall-process}
\end{figure}

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.9\linewidth]{images/grounding.png}
  \caption{The three main components of our architecture and their connections. Submodules are presented in subsequent sections.}
  \label{fig|main-components}
\end{figure}

@@

\begin{figure}[!ht]
\centering
  \includegraphics[width=0.8\linewidth]{images/oro_architecture.png}
  \caption{Overview of the ORO architecture.}
  \label{fig|oro-overview}
\end{figure}

@@

\item \emph{Sees}: An important ability to know about an agent is to predict
``what it can see'', \ie what is within its field of view (FOV). A robot being
able to compute this information can then act accordingly. An example
would be a clarification scenario where the human is searching for an object
and the robot is able to infer that he/she is looking for the one that is not
visible (otherwise the user would not be searching for it).
In Figure~\ref{fig::vis} the field of view of a person is illustrated with a
grey cone (broader one). While he is able to see the two small boxes on the
table in front of him, the big box on his right is out of his FOV, and
therefore, he is not able to see it.

\item \emph{Looks At}: this relation corresponds to what the agent is focused
on, \ie where its focus of attention is directed. This model is based on a
narrower field of view, the field of attention (FOA). Figure~\ref{fig::vis}
shows the field of attention of a person with a green cone (narrower one). In
this example only the grey box satisfies the \concept{looksAt} relation.

\begin{figure}%[!t] 
\centering
\includegraphics[width=\columnwidth]{images/looks.jpg} \caption{Representation
of the field of view (FOV) and the field of attention (FOA) of the human.}
\label{fig::vis} \end{figure}

@@

\item \emph{Reachable}: it allows the robot to estimate the agent's capability
to reach an object, which is fundamental for task planning. For example, if the
user asks the robot to give him/her an object, the robot must compute a transfer
point where the user is able to get the object afterward. 
Figure~\ref{fig::reach} shows different reachability postures for each object
on the table. In the example, the bottle and the box are both reachable for the
human, but the teddy bear is too far. Instead, from the robot's perspective,
the teddy bear is reachable for it, while the bottle is not.

\begin{figure}%[!t] 
	\centering
	\includegraphics[width=0.8\columnwidth]{images/reach.jpg} 
	\caption{Different reaching postures for the human: the bottle and the 
	box are reachable, while the teddy bear is not.} 
	\label{fig::reach} 
\end{figure}

@@

@@

\item \emph{Location according to an object}: We can also refer to object
locations with respect to other objects in the environment, such as \emph{above,
next to, in}, etc. In this work we compute three main relations based on the
bounding box and center of mass of the objects (Figure~\ref{fig::sprelation}):

@@

\begin{figure} 
	\centering
	\includegraphics[width=0.8\columnwidth]{images/spatial_relation.png}
	\caption{Spatial relations between two objects: A) \concept{isOn} relation, 
	B) \concept{isIn} relation, and C) \concept{isNextTo} relation.} 
	\label{fig::sprelation} 
\end{figure}

@@

To this end we have developed a dedicated module called {\sc Dialogs} that processes human
input in natural language, grounds the concepts in the robot's knowledge and
eventually translates the discourse in a set of queries or declarative OWL/RDF
statements.

\begin{figure}%[!ht]
\centering
  \includegraphics[width=0.9\linewidth]{images/dialog_module_simple.png}
  \caption{The {\sc Dialogs} module has three main steps: the parsing,
  the interpretation and the verbalization. The interpretation module is
  responsible for both the \emph{resolution} and the semantic content
  \emph{analysis and translation}.} 
  \label{fig|dialog}
\end{figure}

Figure~\ref{fig|dialog} shows the {\sc Dialogs} module architecture. The user's
input is first pre-processed. For instance, \emph{I'm} constructs are expanded
into \emph{I am} and then parsed. The parser is a custom-made, rule-based (\ie
grammar-free) tool that extracts the grammatical structure from the user's
sentence. Figure~\ref{dialog|parser_output} shows an example of the raw output of the parser for a moderately complex sentence.

@@

In order to better understand the overall process of the {\sc Dialogs} module 
and its relation with ORO, we next describe the different steps of the approach 
based on three examples. In these examples we assume that some initial facts 
are present in the knowledge base, both in the robot's own model and in the 
human's model.  Since the robot tries to ground a
human utterance, all queries are sent to the human model in order to interpret
it from the human perspective.

\subsection{Informational Content Extraction}

\begin{figure}
    \centering
	\begin{tabular}{p{7cm}}
	\emph{Initial knowledge model of} \texttt{human\_01}\\
	\hline
    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\

	\vspace{0.5em}
	\emph{Human input}\\
	\hline
	\hspace{0.3cm}``The yellow banana is big!'' \\

	\vspace{0.5em}
	\emph{Generated partial statements}\\
	\hline
	\hspace{0.3cm}\stmt{?obj type Banana} \\
    	\hspace{0.3cm}\stmt{?obj hasColor yellow} \\
    	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\

	\vspace{0.5em}
	\emph{Newly created statements}\\
	\hline
	\hspace{0.3cm}\stmt{banana\_01 hasSize big} \\
	\end{tabular}
\caption{First example of natural language grounding: the nominal group ``the
yellow banana'' is matched with the individual \concept{banana\_01}.
``$\Rightarrow$'' represents the output of the ontology server.}
\label{dialog|ex1}
\end{figure}

@@

\begin{figure}
    \centering
	\begin{tabular}{p{7cm}}
	\emph{Initial knowledge model of} \texttt{human\_01}\\
	\hline
    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
	\end{tabular} \\

	\begin{tabular}{p{7cm}}
	\emph{Human input}\\
	\hline
    	\hspace{0.3cm}``Give me the banana.'' \\
	\end{tabular} \\

	\begin{tabular}{p{7cm}}
	\emph{Generated partial statements}\\
	\hline
    	\hspace{0.3cm}\stmt{?obj type Banana} \\
	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\

	\end{tabular} \\

	\begin{tabular}{p{7cm}}
	\emph{Newly created statements}\\
	\hline
    	\hspace{0.3cm}\stmt{human\_01 desires situation\_a3f74} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 type Give} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 performedBy myself} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 actsOnObject banana\_01} \\
    	\hspace{0.3cm}\stmt{situation\_a3f74 receivedBy human\_01} \\
	\end{tabular}

@@

The scenario for this game (Figure~\ref{fig|spyGameScenario}) consists on a
face-to-face interaction where the human thinks of an object present in the
environment, while the robot queries the human until either discovering the
object or giving up~\cite{Ros2010a}. A categorization example is presented in
Figure~\ref{fig|objectsSpyGame}. The game starts with the human user giving a
first hint (communication is done through a keyboard and screen), allowing the
robot to start the search filtering those objects that fulfill this first
description. Based on this subset, ORO provides a descriptor (or set of
descriptors) that allows maximum discrimination among objects in the subset.
The robot queries the user about the value of the descriptor (or the most
discriminant among the set of descriptors) and with this new information, the
current subset of objects is filtered again. The process is repeated until
either obtaining a single object that fulfills all the descriptor values, or
failing (\ie no object found).

\begin{figure}%[!ht]
\centering
\begin{scriptsize}
\begin{tikzpicture} %[level/.style={sibling distance=30mm/#1},scale=0.8]
	%[edge from parent fork down,
	%every node/.style={fill=black!30,rounded corners},
	%[parent anchor=east,child anchor=west,grow=east,
	%edge from parent/.style={thick,draw}]
	\node {Artifact}
	child {node {Tableware}
		child {node {Bottle}}
		child {node {Cup}}
		}
	child {node {GameObject}}
	child {node {Furniture}
			child {node {Table}}
			child {node {Chair}}
			child {node {Shelf}}};
\end{tikzpicture}
\end{scriptsize}
\caption{Example of object categorization used in the scenario.}
\label{fig|objectsSpyGame}			
\end{figure}

@@

\begin{figure}%[!t]
\centering
  \subfigure[]
  {\includegraphics[width=0.95\linewidth]{images/situationAssessment1.png}\label{a}}
  \subfigure[]
  {\includegraphics[width=0.95\linewidth]{images/situationAssessment2.png}\label{b}}
	\caption{Situation assessment during initialization: a) before placing the
	objects on the table, and b) after placing them. In each image the snapshots
	correspond to: real environment (top-left sub-image); processed image to
	identify the tagged objects (bottom-left sub-image); and the 3D environment
	(right sub-image).}
	\label{fig|sitAssessRoman}
\end{figure}

First Tom enters the room with some of the things they need to pack: a
toolbox and two videos. He leaves one of the video tapes (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place at each step, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.
The initial information in ORO corresponds to:

\begin{center}
    \begin{tabular}{l}
      \stmt{table type Table}\\
      \stmt{cardBoardBox type Box}\\
      \stmt{toolbox type Box}\\
      \stmt{videoTape1 type VideoTape}\\
      \stmt{videoTape1 label "Lord of the Robots"}\\
      \stmt{videoTape2 type VideoTape}\\
      \stmt{videoTape2 label "Jido-E"}\\
      \stmt{Tom type Human}\\
      \stmt{Jerry type Human}
    \end{tabular}
\end{center}
\vspace{0.10cm}
	SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:

\begin{center}
    \begin{tabular}{l}
      \stmt{cardBoardBox isOn table}
    \end{tabular}	             
\end{center}

Tom enters carrying several objects (Figure~\ref{fig|sitAssessRoman}a) and places them around. Then he leaves. (Figure~\ref{fig|sitAssessRoman}b). The following facts are computed and sent to ORO:

\begin{center}
    \begin{tabular}{l}
      \stmt{toolbox isOn table}\\
      \stmt{videoTape1 isOn table}\\
      \stmt{videoTape1 isNextTo toolbox}\\
      \stmt{toolbox isNextTo videoTape1}\\
      \stmt{videoTape2 isIn cardBoardBox}
    \end{tabular}       
\end{center}

@@

Tom enters the room again while carrying a big box (Figure~\ref{fig|vpt}). He
approaches the table and asks Jido to handle him the video tape: ``Jido, can
you give me the video tape''. The \textsc{Dialogs} module queries the ontology to
identify the object the human is referring to:

\begin{center} 
\stmt{?obj type VideoTape} 
\end{center}

There are two video tapes in the scene: one on
the table, and another one inside the cardboard box. Thus, the knowledge 
base returns both:

\begin{center}
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1, videoTape2]}
\end{center}

However, only one is visible for Tom (the one on the
table). Thus, although there is an ambiguity from the robot's perspective
(since it can see both video tapes), based on the perspective of its human
partner it infers that Tom is referring to the video tape on the table, and not
the one inside the box which is not visible from his view. Therefore,
non-visible objects are removed obtaining:

@@

\begin{figure}%[!ht]
  \centering
  \subfigure[]
  {\includegraphics[width=0.9\linewidth]{images/inTheBox1.jpg}\label{a}}
  \subfigure[]
  {\includegraphics[width=0.9\linewidth]{images/inTheBox2.jpg}\label{b}}
\caption{(a) Jerry asks Jido for the content of the box and (b) points at it.}
  \label{fig|pointing}
\end{figure}

Figure~\ref{fig|pointing}a depicts the third situation where Jerry enters the
living room without knowing where Tom had placed the video tapes. So he first
asks Jido: ``What's in the box?''. Before the robot can answer the question it
has to figure out which box Jerry is talking about. Similar to the previous
situation, there are two available boxes:

\begin{center}
\begin{tabular}{l}
\stmt{?obj type Box}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox, toolbox]}
\end{tabular}
\end{center}

However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring
to\footnote{Note that Jerry is using the definite article \emph{the}: the robot
has to determine which box is relevant. {\sc Dialogs} supports other kinds of
quantification (existantial -- \emph{some} --, universal -- \emph{all} -- , explicit cardinality), as expressed by the combination of definite/indefinite and
singular/plural forms of articles.}: ``Which box, the toolbox or the cardboard box?''
Jerry could now simply answer the question. Instead, he decides to point at it
while indicating: ``That box'' (Figure~\ref{fig|pointing}b). The robot's
perception identifies the {\tt cardBoardBox} as being pointed at and looked
at by the human and updates the ontology with this new information using a
rule available in the commonsense ontology:

\begin{center}
\begin{tabular}{l}
\stmt{Jerry pointsAt carboardBox}\\
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\end{tabular}
\end{center}
In the meantime, the \textsc{Dialogs} module is processing the human verbal input. When trying to resolve the reference ``that'' it is able to merge\footnote{Due to synchronization issues, the user should perform the gesture (pointing at) before answering the robot's question and maintain it until the resolution process takes place.} both sources of information, verbal and gestural, to distinguish the box Jerry refers to:
\begin{center}
\begin{tabular}{l}
\stmt{Jerry focusesAt ?obj}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox]}
\end{tabular}
\end{center}

@@

At this point Jerry wants to know where the other video tape is, and that is exactly
what he asks Jido: ``And where is the other video tape?''. In this occasion, the
\textsc{Dialogs} module is able to interpret that Jerry is not referring to the
video tape which they were just talking about, but to the other one:

\begin{center}
\begin{tabular}{l}
\stmt{?obj type VideoTape}\\
\stmt{?obj differentFrom videoTape2}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1]}
\end{tabular}
\end{center}

Since there is only one possible ``other'' video tape (there are only two video tapes in
the scene), it can directly answer Jerry: ``The other video tape is on the table and
next to the toolbox.''

\begin{center}
\begin{tabular}{l}
\stmt{videoTape1 isOn table}\\
\stmt{videoTape1 isNextTo toolbox}
\end{tabular}
\end{center}

@@

@@

Amongst the other shortcomings of our architecture, neither the \emph{domain of
validity} nor the context of a fact are represented in a satisfying way (we do
store some kind of context --the agent's mental model for instance--). This
information is meta-information on the knowledge. While the ORO framework
allows them through \emph{statement reification}, it does not offer yet a
convenient way to store them. One obvious limitation that derives from the lack
of efficient meta-knowledge is the absence of knowledge history.  With ORO, the
robot always lives in the present.

>>>>

In this work we propose an approach aiming at grounding the interaction with
users (Figure~\ref{fig|overall-process} illustrates the general context). To this end, we have
developed three distinct, inter-related cognitive components:

@@

\begin{figure}[!t]
\centering
  \includegraphics[width=0.8\linewidth]{images/grounding_robot.png}
  \caption{Generic model of cognitive abilities interactions for grounding}
  \label{fig|overall-process}
\end{figure}

%\begin{figure}[!ht]
%\centering
%  \includegraphics[width=0.9\linewidth]{images/grounding.png}
%  \caption{The three main components of our architecture and their connections. Submodules are presented in subsequent sections.}
%  \label{fig|main-components}
%\end{figure}

@@

\begin{figure}[!t]
\centering
  \includegraphics[width=0.8\linewidth]{images/oro_architecture.png}
  \caption{Overview of the ORO architecture.}
  \label{fig|oro-overview}
\end{figure}

@@

\item \emph{Sees}: An important ability to know about an agent is to predict
``what it can see'', \ie what is within its field of view (FOV). A robot being
able to compute this information can then act accordingly. An example
would be a clarification scenario where the human is searching for an object
and the robot is able to infer that he/she is looking for the one that is not
visible (otherwise the user would not be searching for it).
In Figure~\ref{fig::sparkRepresentations}\emph{a} the field of view of a person is illustrated with a
grey cone (broader one). While he is able to see the two small boxes on the
table in front of him, the big box on his right is out of his FOV, and
therefore, he is not able to see it.

\item \emph{Looks At}: this relation corresponds to what the agent is focused
on, \ie where its focus of attention is directed. This model is based on a
narrower field of view, the field of attention (FOA). 
Figure~\ref{fig::sparkRepresentations}\emph{a}
shows the field of attention of a person with a green cone (narrower one). In
this example only the grey box satisfies the \concept{looksAt} relation.

%\begin{figure}%[!t] 
%\centering
%\includegraphics[width=0.8\columnwidth]{images/looks.jpg} \caption{Representation
%of the field of view (FOV) and the field of attention (FOA) of the human.}
%\label{fig::vis} 
%\end{figure}

@@

\item \emph{Reachable}: it allows the robot to estimate the agent's capability
to reach an object, which is fundamental for task planning. For example, if the
user asks the robot to give him/her an object, the robot must compute a transfer
point where the user is able to get the object afterward. 
Figure~\ref{fig::sparkRepresentations}\emph{b} shows different reachability postures for each object
on the table. In the example, the bottle and the box are both reachable for the
human, but the teddy bear is too far. Instead, from the robot's perspective,
the teddy bear is reachable for it, while the bottle is not.

%\begin{figure}%[!t] 
%	\centering
%	\includegraphics[width=0.8\columnwidth]{images/reach.jpg} 
%	\caption{Different reaching postures for the human: the bottle and the 
%	box are reachable, while the teddy bear is not.} 
%	\label{fig::reach} 
%\end{figure}

@@

\begin{figure*}[!t]
	\subfigure[]
	{\includegraphics[width=0.35\linewidth]{images/looks.jpg} \label{(a)}} \quad
	\subfigure[]
	{\includegraphics[width=0.3\linewidth]{images/reach.jpg} \label{(b)}}\qquad	
	\subfigure[]
	{\includegraphics[width=0.25\linewidth]{images/spatial_relation.png} \label{(c)}}

	\caption{(a) Field of view (FOV) and the field of attention (FOA) of the human. (b) Different reaching postures for the human. (c) Spatial relations between two objects: A) \concept{isOn} relation, B) \concept{isIn} relation, and C) \concept{isNextTo} relation.}
	\label{fig::sparkRepresentations}
\end{figure*}

@@

\item \emph{Location according to an object}: We can also refer to object
locations with respect to other objects in the environment, such as \emph{above,
next to, in}, etc. In this work we compute three main relations based on the
bounding box and center of mass of the objects 
(Figure~\ref{fig::sparkRepresentations}\emph{c}):

@@

%\begin{figure} 
%	\centering
%	\includegraphics[width=0.6\columnwidth]{images/spatial_relation.png}
%	\caption{Spatial relations between two objects: A) \concept{isOn} relation, 
%	B) \concept{isIn} relation, and C) \concept{isNextTo} relation.} 
%	\label{fig::sprelation} 
%\end{figure}

@@

\begin{figure}[!t]
\centering
  \includegraphics[width=0.9\linewidth]{images/dialog_module_simple.png}
  \caption{The {\sc Dialogs} module has three main steps: the parsing,
  the interpretation and the verbalization. The interpretation module is
  responsible for both the \emph{resolution} and the semantic content
  \emph{analysis and translation}.} 
  \label{fig|dialog}
\end{figure}

To this end we have developed a dedicated module called {\sc Dialogs} that processes human
input in natural language, grounds the concepts in the robot's knowledge and
eventually translates the discourse in a set of queries or declarative OWL/RDF
statements. Figure~\ref{fig|dialog} shows the {\sc Dialogs} module architecture. The user's
input is first pre-processed. For instance, \emph{I'm} constructs are expanded
into \emph{I am} and then parsed. The parser is a custom-made, rule-based (\ie
grammar-free) tool that extracts the grammatical structure from the user's
sentence. Figure~\ref{dialog|parser_output} shows an example of the raw output of the parser for a moderately complex sentence.

@@

In order to better understand the overall process of the {\sc Dialogs} module 
and its relation with ORO, we next describe the different steps of the approach 
based on three examples. In these examples we assume that some initial facts 
are present in the knowledge base (section~\ref{modeling_real_world} discusses how 
the initial knowledge can be acquired), both in the robot's own model and in the human's model.  Since the robot tries to ground a human utterance, all queries are sent to the human model in order to interpret it from the human perspective.

\subsection{Informational Content Extraction}
\label{informational_content_extraction}

%\begin{figure}
%    \centering
%	\begin{tabular}{p{7cm}}
%	\emph{Initial knowledge model of} \texttt{human\_01}\\
%	\hline
%    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
%    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
%	
%	\vspace{0.5em}
%	\emph{Human input}\\
%	\hline
%	\hspace{0.3cm}``The yellow banana is big!'' \\
%
%	\vspace{0.5em}
%	\emph{Generated partial statements}\\
%	\hline
%	\hspace{0.3cm}\stmt{?obj type Banana} \\
%    	\hspace{0.3cm}\stmt{?obj hasColor yellow} \\
%    	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
%
%	\vspace{0.5em}
%	\emph{Newly created statements}\\
%	\hline
%	\hspace{0.3cm}\stmt{banana\_01 hasSize big} \\
%	\end{tabular}
%\caption{First example of natural language grounding: the nominal group ``the
%yellow banana'' is matched with the individual \concept{banana\_01}.
%``$\Rightarrow$'' represents the output of the ontology server.}
%\label{dialog|ex1}
%\end{figure}

\begin{figure}
    \centering
%	\begin{tabular}{p{0.5\columnwidth} | p{0.5\columnwidth}}}
	\begin{tabular}{l|l}
	\emph{Initial knowledge} \texttt{human\_01} &
	\emph{Human input}\\

	\hline

    	\stmt{banana\_01 type Banana} &
	``The yellow banana is big!'' \\

    	\stmt{banana\_01 hasColor yellow} & \\
	\vspace{0.5em}\\
	\hline

	\emph{Generated partial statements} &
	\emph{Newly created statements}\\
	\hline

	\stmt{?obj type Banana} &
	\hspace{0.2cm}\stmt{banana\_01 hasSize big} \\

    	\stmt{?obj hasColor yellow} & \\
    	\hspace{0.2cm}$\Rightarrow$ \concept{?obj = banana\_01}\\

	\hline
	\end{tabular}
\caption{First example: content extraction.
``$\Rightarrow$'' represents the output of the ontology server.}
\label{dialog|ex1}
\end{figure}

@@

%\begin{figure}
%    \centering
%	\begin{tabular}{p{7cm}}
%	\emph{Initial knowledge model of} \texttt{human\_01}\\
%	\hline
%    	\hspace{0.3cm}\stmt{banana\_01 type Banana} \\
%    	\hspace{0.3cm}\stmt{banana\_01 hasColor yellow}\\
%	\end{tabular} \\
%
%	\vspace{0.5em}
%
%	\begin{tabular}{p{7cm}}
%	\emph{Human input}\\
%	\hline
%    	\hspace{0.3cm}``Give me the banana.'' \\
%	\end{tabular} \\
%
%	\vspace{0.5em}
%
%	\begin{tabular}{p{7cm}}
%	\emph{Generated partial statements}\\
%	\hline
%    	\hspace{0.3cm}\stmt{?obj type Banana} \\
%	\hspace{0.7cm}$\Rightarrow$ \concept{?obj = banana\_01}\\
%
%	\end{tabular} \\
%
%	\vspace{0.5em}
%
%	\begin{tabular}{p{7cm}}
%	\emph{Newly created statements}\\
%	\hline
%    	\hspace{0.3cm}\stmt{human\_01 desires situation\_a3f74} \\
%    	\hspace{0.3cm}\stmt{situation\_a3f74 type Give} \\
%    	\hspace{0.3cm}\stmt{situation\_a3f74 performedBy myself} \\
%    	\hspace{0.3cm}\stmt{situation\_a3f74 actsOnObject banana\_01} \\
%    	\hspace{0.3cm}\stmt{situation\_a3f74 receivedBy human\_01} \\
%	\end{tabular}
%
%\caption{Second example: processing an order.}
%\label{dialog|ex2}
%\end{figure}

\begin{figure}
    \centering
	\begin{tabular}{l|l}
	\emph{Initial knowledge} \texttt{human\_01} &
	\emph{Human input}\\

	\hline

    	\stmt{banana\_01 type Banana} &
	``Give me the banana.'' \\

    	\stmt{banana\_01 hasColor yellow} & \\
	\vspace{0.5em}\\
	\hline

	\emph{Generated partial statements} &
	\emph{Newly created statements}\\
	\hline
    	\stmt{?obj type Banana} & 
	\stmt{human\_01 desires sit\_a3} \\

	\hspace{0.2cm}$\Rightarrow$ \concept{?obj = banana\_01}
    	& \stmt{sit\_a3 performedBy myself} \\
    	& \stmt{sit\_a3 actsOnObject banana\_01} \\
    	& \stmt{sit\_a3 receivedBy human\_01} \\
	\end{tabular}

@@

The scenario for this game (Figure~\ref{fig|spyGameScenario}) consists on a
face-to-face interaction where the human thinks of an object present in the
environment, while the robot queries the human until either discovering the
object or giving up~\cite{Ros2010a}. The game starts with the human user giving a
first hint (communication is done through a keyboard and screen), allowing the
robot to start the search filtering those objects that fulfill this first
description. Based on this subset, ORO provides a descriptor (or set of
descriptors) that allows maximum discrimination among objects in the subset.
The robot queries the user about the value of the descriptor (or the most
discriminant among the set of descriptors) and with this new information, the
current subset of objects is filtered again. The process is repeated until
either obtaining a single object that fulfills all the descriptor values, or
failing (\ie no object found).

%\begin{figure}%[!ht]
%\centering
%\begin{scriptsize}
%\begin{tikzpicture} %[level/.style={sibling distance=30mm/#1},scale=0.8]
%	%[edge from parent fork down,
%	%every node/.style={fill=black!30,rounded corners},
%	%[parent anchor=east,child anchor=west,grow=east,
%	%edge from parent/.style={thick,draw}]
%	\node {Artifact}
%	child {node {Tableware}
%		child {node {Bottle}}
%		child {node {Cup}}
%		}
%	child {node {GameObject}}
%	child {node {Furniture}
%			child {node {Table}}
%			child {node {Chair}}
%			child {node {Shelf}}};
%\end{tikzpicture}
%\end{scriptsize}
%\caption{Example of object categorization used in the scenario.}
%\label{fig|objectsSpyGame}			
%\end{figure}

@@

\begin{figure*}
\centering
	\subfigure[]
	{\includegraphics[width=0.3\linewidth]{images/situationAssessment1.png}\label{(a)}}
	\subfigure[]
	{\includegraphics[width=0.3\linewidth]{images/situationAssessment2.png}\label{(b)}}
	\subfigure[]
  	{\includegraphics[width=0.3\linewidth]{images/inTheBox2.jpg}\label{c}}
	\caption{Situation assessment during initialization (before (a) and after (b) placing the objects on the table). In each image the snapshots correspond to: real environment (top-left sub-image); processed image to identify the tagged objects (bottom-left sub-image); and the 3D environment (right sub-image).(c) Jerry asks Jido for the content of the box while pointing at it.}
	\label{fig|movingToLondon}
\end{figure*}

%\begin{figure}%[!t]
%\centering
%  \subfigure[]
%  {\includegraphics[width=0.95\linewidth]{images/situationAssessment1.png}\label{a}}
%  \subfigure[]
%  {\includegraphics[width=0.95\linewidth]{images/situationAssessment2.png}\label{b}}
%	\caption{Situation assessment during initialization: a) before placing the
%	objects on the table, and b) after placing them. In each image the snapshots
%	correspond to: real environment (top-left sub-image); processed image to
%	identify the tagged objects (bottom-left sub-image); and the 3D environment
%	(right sub-image).}
%	\label{fig|sitAssessRoman}
%\end{figure}

First Tom enters the room with some of the things they need to pack: a
toolbox and two videos. He leaves one of the video tapes (``Jido-E'') inside one of
the boxes (the cardboard box), and the other one (``Lord of the Robots'') on the
table. We next describe how the situation assessment takes place at each step, \ie how the
ontology is updated with the information obtained from the geometric reasoner
SPARK.
The initial information in ORO corresponds to:
\begin{center}
    \begin{tabular}{ll}
      \stmt{table type Table} & \stmt{Tom type Human}\\
      \stmt{cardBoardBox type Box} & \stmt{Jerry type Human}\\
      \stmt{toolbox type Box} & \stmt{videoTape1 type VideoTape}\\
      \stmt{videoTape2 type VideoTape} & \stmt{videoTape2 label "Jido-E"}   
      \end{tabular}
      \stmt{videoTape1 label "Lord of the Robots"}
\end{center}
\vspace{0.10cm}
	SPARK detects that there is a cardboard box on the table. It thus sends the fact to ORO:

\begin{center}
    \begin{tabular}{l}
      \stmt{cardBoardBox isOn table}
    \end{tabular}	             
\end{center}
Tom enters carrying several objects (Figure~\ref{fig|movingToLondon}a) and places them around. Then he leaves. (Figure~\ref{fig|movingToLondon}b). The following facts are computed and sent to ORO:
\begin{center}
    \begin{tabular}{ll}
      \stmt{toolbox isOn table} & \stmt{toolbox isNextTo videoTape1}\\
      \stmt{videoTape1 isOn table} & \stmt{videoTape2 isIn cardBoardBox}
      \end{tabular}
      \stmt{videoTape1 isNextTo toolbox}
\end{center}

@@

Tom enters the room again while carrying a big box (Figure~\ref{fig|vpt}). He
approaches the table and asks Jido to handle him the video tape: ``Jido, can
you give me the video tape''. The \textsc{Dialogs} module queries the ontology to
identify the object the human is referring to:
\begin{center} 
\stmt{?obj type VideoTape} 
\end{center}
There are two video tapes in the scene: one on
the table, and another one inside the cardboard box. Thus, the knowledge 
base returns both: 
\begin{center}
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1, videoTape2]}
\end{center}
However, only one is visible for Tom (the one on the
table). Thus, although there is an ambiguity from the robot's perspective
(since it can see both video tapes), based on the perspective of its human
partner it infers that Tom is referring to the video tape on the table, and not
the one inside the box which is not visible from his view. Therefore,
non-visible objects are removed obtaining:

@@

%\begin{figure}%[!ht]
%  \centering
%  \subfigure[]
%  {\includegraphics[width=0.9\linewidth]{images/inTheBox1.jpg}\label{a}}
%  \subfigure[]
%  {\includegraphics[width=0.9\linewidth]{images/inTheBox2.jpg}\label{b}}
%\caption{(a) Jerry asks Jido for the content of the box and (b) points at it.}
%  \label{fig|pointing}
%\end{figure}

In this last situation Jerry enters the
living room without knowing where Tom had placed the video tapes. So he first
asks Jido: ``What's in the box?''. Before the robot can answer the question it
has to figure out which box Jerry is talking about. Similar to the previous
situation, there are two available boxes: 
\begin{center}
\begin{tabular}{l}
\stmt{?obj type Box}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox, toolbox]}
\end{tabular}
\end{center}
However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring
to\footnote{Note that Jerry is using the definite article \emph{the}: the robot
has to determine which box is relevant. {\sc Dialogs} supports other kinds of
quantification (existantial -- \emph{some} --, universal -- \emph{all} -- , explicit cardinality), as expressed by the combination of definite/indefinite and
singular/plural forms of articles.}: ``Which box, the toolbox or the cardboard box?''
Jerry could now simply answer the question. Instead, he decides to point at it
while indicating: ``That box'' (Figure~\ref{fig|movingToLondon}c). The robot's
perception identifies the {\tt cardBoardBox} as being pointed at and looked
at by the human and updates the ontology with this new information using a
rule available in the commonsense ontology:
\begin{center}
\begin{tabular}{r}
\stmt{Jerry pointsAt carboardBox}, 
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\end{tabular}
\end{center}
In the meantime, the \textsc{Dialogs} module is processing the human verbal input. When trying to resolve the reference ``that'' it is able to merge\footnote{Due to synchronization issues, the user should perform the gesture (pointing at) before answering the robot's question and maintain it until the resolution process takes place.} both sources of information, verbal and gestural, to distinguish the box Jerry refers to:
\begin{center}
\begin{tabular}{l}
\stmt{Jerry focusesAt ?obj}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [cardBoardBox]}
\end{tabular}
\end{center}

@@

At this point Jerry wants to know where the other video tape is, and that is exactly
what he asks Jido: ``And where is the other video tape?''. In this occasion, the
\textsc{Dialogs} module is able to interpret that Jerry is not referring to the
video tape which they were just talking about, but to the other one:
\begin{center}
\begin{tabular}{l}
\stmt{?obj type VideoTape}\\
\stmt{?obj differentFrom videoTape2}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1]}
\end{tabular}
\end{center}

Since there is only one possible ``other'' video tape (there are only two video tapes in
the scene), it can directly answer Jerry: ``The other video tape is on the table and
next to the toolbox.''
\begin{center}
\begin{tabular}{l}
\stmt{videoTape1 isOn table}, 
\stmt{videoTape1 isNextTo toolbox}
\end{tabular}
\end{center}

@@

Note also that perception issues have been solved in our experiments by using a tag-based object identification
method. In section~\ref{informational_content_extraction} we give an example
where the human says ``the yellow banana is big''. It is assumed in the example
that the robot already knows about a banana instance that is yellow. In our
experiments, this kind of knowledge was either hard coded in scenario-specific
ontologies (\eg \stmt{banana\_01 type Banana} where \concept{banana\_01}
is the id of the banana's tag) or taught to the robot with prescriptive
sentences like ``Learn that this is a banana'' while pointing at the banana's
tag. It would be interesting to extend this approach with automatic classifiers
(for colour, size, etc.). If the robot later discovers a yellowish and large
object, an utterance like ``the yellow banana is big'' could be used to assert
that this object is a banana.  A similar approach focused on the combination of visual perception and communication modalities to achieve visual learning has been 
developed by~\cite{Vrecko2009}.

@@

Amongst the other shortcomings of our architecture, neither the \emph{domain of
validity} nor the context of a fact are represented in a satisfying way (we do
store some kind of context --the agent's mental model for instance). This
information is meta-information on the knowledge. While the ORO framework
allows them through \emph{statement reification}, it does not offer yet a
convenient way to store them. One obvious limitation that derives from the lack
of efficient meta-knowledge is the absence of knowledge history.  With ORO, the
robot always lives in the present.
